---
title: "Analysis"
output:
  word_document: default
  pdf_document: default
---

<style type="text/css">
  body{
  font-size: 14pt;
}
</style>


```{r}
library(corrplot)
library(RColorBrewer)
library(readxl)
```


```{r}
data = read_excel("housing.xls")
head(data)
```

```{r}
str(data)

```
most columns are numerical variables

```{r}
summary(data)
```
here we find a summary of our dataset
```{r}
colSums(is.na(data))
```
we observe that there is no missing values in our dataset



```{r}
M <-cor(data)
corrplot(M, order="hclust",
         col=brewer.pal(n=8, name="RdYlBu"))
```


we first notice that median value of homes is positively correlated with number of rooms per dwelling which is obvious.

we notice that lower status is positively correlated with crime rates and negatively correlated with home values
and nitric oxide concentration which is emitted mostly by vehicles and industrial equiment is positively correlated to percentage of lower status.

from the aforementioned we conclude that the population near the industrial zone tend to be lower status which feeds more into the crime rate.

moreover student to teacher ratio indicated by ptratio is negatively correlated with home values which makes sense since the less teachers available for students in a specific town the less valuable the homes in that town will be due to lack of education and eventually lower status.

we will explore these findings further in later visualizations to validate our analysis.


```{r}
plot(MEDV~RM, data)

```
```{r}
plot(MEDV~LSTAT, data)
```
```{r}
plot(CRIM~LSTAT, data)
```

```{r}
plot(NOX~LSTAT, data)
```
```{r}
plot(MEDV~PTRATIO, data)
```
```{r}
plot(data[,c(1,7,6,12,13,14)],pch=3)
```

These plots confirm our findings which we restate :

* median value of homes is positively correlated with number of rooms per dwelling.
* lower status is positively correlated with crime rates and negatively correlated with home values.
* nitric oxide concentration which is emitted mostly by vehicles and industrial equiment is positively correlated to percentage of lower status.
* population near the industrial zone tend to be lower status which feeds more into the crime rate.
* student to teacher ratio indicated by ptratio is negatively correlated with home values which makes sense since the less teachers available for students in a specific town the less valuable the homes in that town will be due to lack of education and eventually lower status.
