---
title: "Classification"
output:
  word_document: default
  pdf_document: default
---

<style type="text/css">
  body{
  font-size: 14pt;
}
</style>

```{r}
library(readxl)
library(randomForest)
library(caTools)
library(caret)
library(e1071)
```

## Data Preparation

```{r}
data = read_excel("housing.xls")
head(data)
```
We see that our target variable is balanced.

```{r}
barplot(table(data$CR01))
```

We omit the column CRIM because it leaks information about our target variable CR01 since they both are indicators of crime rate. (if we include it we get 100% accuracy because our model will everything it needs from that column)

```{r}
data = subset(data, select = -c(CRIM) )
```

we partition our data into train and test datasets so we can benchmark the performance of our model
```{r}
set.seed(12345)

data$CR01 <- as.factor(data$CR01)

#Do data partitioning
inTrain <- createDataPartition(y = data$CR01, p = 0.70, list = FALSE)
training <- data[inTrain,]
testing <- data[-inTrain,]
```

## SVM

We start off with an SVM as a classifier and we choose our features from our correlation matrix in the analysis notebook, We then predict on the test dataset and retrieve all the relevant metrics and statistics like accuracy, confusion matrix

```{r}
set.seed(12345)
fit.svm <- svm(formula = CR01 ~ ZN + NOX + AGE + DIS + RAD + TAX + PTRATIO + LSTAT + MEDV, data = training, type = 'C-classification', kernel = 'radial')

set.seed(12345)
pred.svm <- predict(fit.svm, newdata = testing)

```
we get an accuracy of 0.92 which is good but we see from the confusion matrix that we misclassify relatively more areas that have higher crime rate which can be a problem.

## Evaluation

```{r}
print(fit.svm)
confusionMatrix(pred.svm, testing$CR01)
summary(fit.svm)
```



## Random Forest

We train a random forest model on all the features since the randomForest package does feature selection by default, We then predict on the test dataset and retrieve all the relevant metrics and statistics like accuracy, confusion matrix and feature importances

```{r}
set.seed(12345)
fit.rf <- fit.rf <- randomForest(formula = CR01~., data = training)

set.seed(12345)
pred.rf <- predict(fit.rf, newdata = testing)

```


## Evaluation

```{r}

print(fit.rf)

confusionMatrix(pred.rf, testing$CR01)

fit.rf$importance
```
We get an accuracy of 0.9267 which is already very good, we also find some features that our model didn't really use
like ZN, CHAS, LSTAT and RM
